---
title: "1- data organizing- prelim full model"
output: html_document
---
---
title: "Initial Full Model"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: html_document
---

Here is our first pass at running the full model. I included most of the code in the output just so everyone can see exactly what is happening under the hood (and maybe catch any mistakes and/or flaws in logic on my part). 
```{r setup, include=FALSE, message=FALSE, warning=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("tidyverse")
# install.packages("readr")
# install.packages("here")
# install.packages("tmle")
# install.packages("gbm")
# install.packages("earth")
# install.packages("gam")
# install.packages("arm")
# install.packages("stats")
# install.packages("ltmle")

library(readr)
library(here)
library(tmle)
library(gbm)
library(earth)
library(gam)
library(arm)
library(stats)
library(ltmle)
library(tidyverse)
expit<-function(p){
  exp(p)/(1+exp(p))
}
```


```{r load files, include=FALSE, message=FALSE, warning=FALSE, error=FALSE}
dat <- read_csv(here("data", "partial_set_covonly_091719_plus_demo.csv"))

# deal with some extreme outlier values that are probably messing up our estimates
dat <- dat %>% 
  mutate(
    mn_bed_cut = if_else(mn_bed > 9, 9, mn_bed, mn_bed), 
    mn_bed_perunit  = mn_bed / ct_units,
    mn_bath_cut = if_else(mn_bath > 9, 9, mn_bath, mn_bath), 
    mn_bath_perunit  = mn_bath / ct_units,
    med_sale_cut = if_else(med_sale_res_5 >= 1000000,
                                 1000000, med_sale_res_5,
                                 med_sale_res_5), 
    med_sale_perunit = med_sale_res_5 / ct_units, 

  )

```

##### Recoding
Some recoding to address missing values. 
```{r, message=FALSE, warning=FALSE, error=FALSE}
dat2 <- dat %>% 
  dplyr::select(-(`freq_`)) %>% 
  mutate_at(vars(ct_par:tot_trans), ~if_else(is.na(.), 0, .))

# for now, assuming missing any_liv is the same as any_liv==0
dat3 <- dat2 %>% 
  mutate(
    mn_bed_perunit = case_when(is.na(mn_bed_perunit) & any_liv==1 ~ -999, 
                       is.na(mn_bed_perunit) & is.na(any_liv) ~ -888,
                       is.na(mn_bed_perunit) & any_liv == 0 ~ -888, 
                       TRUE ~ mn_bed_perunit), 
    bed888 = if_else(mn_bed_perunit == -888, 0, 1), 
    bed999 = if_else(mn_bed_perunit == -999, 0, 1), 
    
    mn_bath_perunit = case_when(is.na(mn_bath_perunit) & any_liv==1 ~ -999, 
                        is.na(mn_bath_perunit) & is.na(any_liv) ~ -888,
                        is.na(mn_bath_perunit) & any_liv == 0 ~ -888, 
                        TRUE ~ mn_bath_perunit), 
    bath888 = if_else(mn_bath_perunit == -888, 0, 1), 
    bath999 = if_else(mn_bath_perunit == -999, 0, 1), 
    
    per_brick = case_when(is.na(per_brick) & any_liv==1 ~ -999, 
                          is.na(per_brick) & is.na(any_liv) ~ -888,
                          is.na(per_brick) & any_liv == 0 ~ -888, 
                          TRUE ~ per_brick), 
    brick888 = if_else(per_brick == -888, 0, 1), 
    brick999 = if_else(per_brick == -999, 0, 1), 
    
    per_singfam = case_when(is.na(per_singfam) & any_liv==1 ~ -999, 
                            is.na(per_singfam) & is.na(any_liv) ~ -888,
                            is.na(per_singfam) & any_liv == 0 ~ -888, 
                            TRUE ~ per_singfam), 
    singfam888 = if_else(per_singfam == -888, 0, 1), 
    singfam999 = if_else(per_singfam == -999, 0, 1),
    
    per_com = case_when(is.na(per_com) & any_liv==1 ~ -999, 
                        is.na(per_com) & is.na(any_liv) ~ -888,
                        is.na(per_com) & any_liv == 0 ~ -888, 
                        TRUE ~ per_com), 
    com888 = if_else(per_com == -888, 0, 1), 
    com999 = if_else(per_com == -999, 0, 1),
    
    per_ind = case_when(is.na(per_ind) & any_liv==1 ~ -999, 
                        is.na(per_ind) & is.na(any_liv) ~ -888,
                        is.na(per_ind) & any_liv == 0 ~ -888, 
                        TRUE ~ per_ind), 
    ind888 = if_else(per_ind == -888, 0, 1), 
    ind999 = if_else(per_ind == -999, 0, 1),
    
    per_vc_lot = case_when(is.na(per_vc_lot) & any_liv==1 ~ -999, 
                           is.na(per_vc_lot) & is.na(any_liv) ~ -888,
                           is.na(per_vc_lot) & any_liv == 0 ~ -888, 
                           TRUE ~ per_vc_lot), 
    vc_lot888 = if_else(per_vc_lot == -888, 0, 1), 
    vc_lot999 = if_else(per_vc_lot == -999, 0, 1), 
    
    med_age = case_when(is.na(med_age) & any_bldg == 1 ~ -999, 
                        is.na(med_age) & is.na(any_bldg) ~ -888, 
                        TRUE ~ med_age), 
    med_age888 = if_else(med_age == -888, 0, 1), 
    med_age999 = if_else(med_age == -999, 0, 1), 
    
    med_sale_perunit = case_when(is.na(med_sale_perunit) ~ -888, 
                               TRUE ~ med_sale_perunit), 
    med_sale888 = if_else(med_sale_perunit == -888, 0, 1))






vc <- dat3 %>% 
  select(hex_1000, year, pt1violence_densl2)

vc2 <- vc %>% 
  pivot_wider(values_from = pt1violence_densl2, 
              names_from = year, 
              names_prefix = "pt1v_dens")

dat3 <- dat3 %>% 
  left_join(vc2)

rm(vc, vc2)




```

##### TMLE
```{r, include=FALSE, message=FALSE, warning=FALSE, error=FALSE}

# interact all acs variables with anypop flag
acs <- names(dat3[64:77])
nn <- 0
for (i in seq_along(acs)){
  nn[i] <-paste0(acs[i], ":", "anypop")
}

acs2 <- str_c(nn, collapse =   " + ")

```

Put together the models: 

* Treatment model: predict the probability of demolition in 2011 or 2012, given 2010 hexagon attributes
    * note: demolition variables coded to indicate demolition in the previous year
* Outcome model: estimate the number of part 1 violent crimes that would have occurred per hexagon in 2013, had there been no 2011-2012 demolition
    * the outcome variable, pt1violence, is the count of all Violent Crime Index crimes (i.e., with and without a firearm)

```{r, message=FALSE, warning=FALSE, error=FALSE}


dat4 <- dat3 %>% 
  mutate(
    demo1112 = if_else( demo2012 > 0 | demo2013 > 0, 1, 0), 
    demo1213 = if_else( demo2013 > 0 | demo2014 > 0, 1, 0), 
    demo1314 = if_else( demo2014 > 0 | demo2015 > 0, 1, 0), 
    demo1415 = if_else( demo2015 > 0 | demo2016 > 0, 1, 0)
    
    )

# drop hex_1000_unk- the bucket for all crimes of unknown location
dat4 <- dat4 %>% 
  filter(hex_1000 != "hex_1000_unk")


# break into 2 sets of datasets (all hexagons, all hexagons with >0 parcels)- 
# hex10 for baseline characteristics
# hex13 for outcome statistics
           
# all 2958 hexagons
hex10 <- dat4 %>% 
  filter(year == 2010)

# the 2624 hexagons with >0 parcels
hex10x <- dat4 %>% 
  filter(year == 2010, any_parcels == 1)


hex13 <- dat4 %>% 
  filter(year == 2013)
hex13x <- dat4 %>% 
  filter(year == 2013, any_parcels == 1)

txmodel <- "demo1112 ~ mn_bed_perunit:bed888:bed999 + mn_bath_perunit:bath888:bath999 + per_brick:brick888:brick999 + per_singfam:singfam888:singfam999 + per_com:com888:com999 + per_ind:ind888:ind999 +  per_vc_lot:vc_lot888:vc_lot999 + med_age:med_age888:med_age999  + occ_manager:anypop + high:anypop + bs:anypop + pov:anypop + hh_pubasst:anypop + male:anypop + nhb:anypop + hisp:anypop + pop18to29:anypop + popls18:anypop + pop65p:anypop + single_hh:anypop + unemp:anypop + medinc:anypop + med_sale_perunit:med_sale888 + offprem_dens + pt1violence + pt1v_dens2010"



```





##### Model- all hexagons (n = 2957)

###### estimate propensity score- predicted probability of any demolition in 2010-2012, given 2010 attributes

```{r, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
hex13$pscore <- predict(glm(formula=txmodel, data=hex10, family="binomial"), type="response")
#practical positivity violations
summary(hex13$pscore)

```


```{r, message=FALSE, warning=FALSE, error=FALSE}

#Create data set for ltmle. 
#Order: W,A,Y
datl <- hex13 %>%
  dplyr::select(
    c(
      mn_bed_perunit,
      bed888,
      bed999,
      mn_bath_perunit,
      bath888,
      bath999,
      per_brick,
      brick888,
      brick999,
      per_singfam,
      singfam888,
      singfam999,
      per_com,
      com888,
      com999,
      per_ind,
      ind888,
      ind999,
      per_vc_lot,
      vc_lot888,
      vc_lot999,
      med_age,
      med_age888,
      med_age999,
      occ_manager,
      anypop,
      high,
      bs,
      pov,
      hh_pubasst,
      male,
      nhb,
      hisp,
      pop18to29,
      popls18,
      pop65p,
      single_hh,
      unemp,
      medinc,
      offprem_dens,
      med_sale_perunit,
      med_sale888,
      pt1v_dens2010,
      pt1v_dens2011,
      pt1v_dens2012,
      demo1112,
      pt1violence
    )
  )



                
#Run TMLE supplying weights
gform <- array(c(hex13$pscore, 1 - hex13$pscore), dim = c(length(hex13$pscore), 1, 2))
    
#Run tmle
tmlesupplyingwtsl<-ltmle(data = datl, Anodes = "demo1112", Lnodes = NULL, Ynodes = "pt1violence", Qform=c(pt1violence="Q.kplus1 ~ demo1112 + mn_bed_perunit:bed888:bed999 + mn_bath_perunit:bath888:bath999 + per_brick:brick888:brick999 + per_singfam:singfam888:singfam999 + per_com:com888:com999 + per_ind:ind888:ind999 +  per_vc_lot:vc_lot888:vc_lot999 + med_age:med_age888:med_age999  + occ_manager:anypop + high:anypop + bs:anypop + pov:anypop + hh_pubasst:anypop + male:anypop + nhb:anypop + hisp:anypop + pop18to29:anypop + popls18:anypop + pop65p:anypop + single_hh:anypop + unemp:anypop + medinc:anypop + offprem_dens + med_sale_perunit:med_sale888 + pt1v_dens2010 + pt1v_dens2011 + pt1v_dens2012"),gform = gform, abar = list(1,0))

summary(tmlesupplyingwtsl)

```


```{r}
mean(hex13$pt1violence)
```

Calculating the difference between what we observed (1.884004 crimes/hexagon) and the number of crimes/hexagon if there was no demolition in any hexagon (Control Estimate: 1.4353), our (current) model suggests that 2011 & 2012 demolitions were responsible for an additional .45 crimes per hexagon.
<br>
<br>


##### Model- only hexagons with parcels (n = 2624)

###### estimate propensity score- predicted probability of any demolition in 2010-2012, given 2010 attributes
```{r, message=FALSE, warning=FALSE, error=FALSE}
#estimate propensity score
hex13x$pscore <- predict(glm(formula=txmodel, data=hex10x, family="binomial"), type="response")
#practical positivity violations
summary(hex13x$pscore)

```


```{r, message=FALSE, warning=FALSE, error=FALSE}

 #Create data set for ltmle. 
 #Order: W,A,Y
datl <- hex13x %>%
  dplyr::select(
    c(
      mn_bed_perunit,
      bed888,
      bed999,
      mn_bath_perunit,
      bath888,
      bath999,
      per_brick,
      brick888,
      brick999,
      per_singfam,
      singfam888,
      singfam999,
      per_com,
      com888,
      com999,
      per_ind,
      ind888,
      ind999,
      per_vc_lot,
      vc_lot888,
      vc_lot999,
      med_age,
      med_age888,
      med_age999,
      occ_manager,
      anypop,
      high,
      bs,
      pov,
      hh_pubasst,
      male,
      nhb,
      hisp,
      pop18to29,
      popls18,
      pop65p,
      single_hh,
      unemp,
      medinc,
      offprem_dens,
      med_sale_perunit,
      med_sale888,
      pt1v_dens2010,
      pt1v_dens2011,
      pt1v_dens2012,
      demo1112,
      pt1violence
    )
  )


                
#Run TMLE supplying weights
gform <- array(c(hex13x$pscore, 1 - hex13x$pscore), dim = c(length(hex13x$pscore), 1, 2))
    
#Run tmle
tmlesupplyingwtsl<-ltmle(data = datl, Anodes = "demo1112", Lnodes = NULL, Ynodes = "pt1violence", Qform=c(pt1violence="Q.kplus1 ~ demo1112 + mn_bed_perunit:bed888:bed999 + mn_bath_perunit:bath888:bath999 + per_brick:brick888:brick999 + per_singfam:singfam888:singfam999 + per_com:com888:com999 + per_ind:ind888:ind999 +  per_vc_lot:vc_lot888:vc_lot999 + med_age:med_age888:med_age999  + occ_manager:anypop + high:anypop + bs:anypop + pov:anypop + hh_pubasst:anypop + male:anypop + nhb:anypop + hisp:anypop + pop18to29:anypop + popls18:anypop + pop65p:anypop + single_hh:anypop + unemp:anypop + medinc:anypop + offprem_dens + med_sale_perunit:med_sale888 + pt1v_dens2010 + pt1v_dens2011 + pt1v_dens2012" ),gform = gform, abar = list(1,0))

summary(tmlesupplyingwtsl)

```

The observed number of Pt.1 Violent crimes in 2013: 

```{r}
mean(hex13x$pt1violence)

```


Our (current) model suggests that 2011 & 2012 demolitions were responsible for an additional .5 crimes per hexagon.
